---
title: Evaporation soil model calibration
layout: post
subtitle: Gonzalo M. Díaz
leafletmap: true
always_allow_html: yes
last_modified_at: 2024-11-10
output: 
  md_document:
    variant: gfm
    preserve_yaml: true
---

Post for calibration of soil model evaporation output

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

First, the aws.wrfsmn library should be open:

```{r}
library("aws.wrfsmn")
```

The example data to use will be ‘eva’ and can be visualize:

```{r}
options(width = 200)
head(eva)
```

The data has several columns with different hydrometeorological variables, for example: precipitation, evaporation, runoff, soil moisture, etc. All these variables are obtained from the VIC (Variable Infiltration Capacity, <https://vic.readthedocs.io/en/master/Overview/ModelOverview/>) soil model, except for the 'evapo_obs' column, which represents the observed data of evaporation for the same location. The data goes from 2015-01-01 to 2017-12-31.

## Definition of predictors variables

The variables chosen for adjustment can be any of those modeled. Here they will be:

```{r}
predictors.variables <- c('OUT_PREC', 'OUT_EVAP', 'OUT_RUNOFF', 'OUT_BASEFLOW',
                          'OUT_SOIL_MOIST_lyr_1', 'OUT_SURF_TEMP')
```

There is no limit on which variables to take, they could be more or less depending on what the user wants.

## Definition of parameters of the Multiple Linear Regression

The data now will be trained with the 2015-01-01 to 2016-12-31 period using 'multiple.guidance' function with the *predictors.variables* vector:

```{r}
data <- eva
data.training <- data[1:which(data$Dates == "2016-12-31"),]

ml.model <- multiple.guidance(input.data = data.training,
                              predictand = 'evapo_obs',
                              predictors = predictors.variables)
ml.model$coefficients
```

Now, the parameters can be used to evaluate the model in any dataset. Here it is applied to the same training period:
```{r}
train.eval <- mg.evaluation(input.data = data.training, predictand = 'evapo_obs',
                            predictors = predictors.variables,
                            var.model = 'OUT_EVAP',
                            lmodel = ml.model)
```

The second element of the list has the statistics parameters of the calibration:
```{r}
train.eval[[2]]
```

And the plot can be visualize with 'ploting' function, but first the monthly data is calculated (for better visualization) with 'daily2monthly' function:
```{r}
ploting(daily2monthly(data = train.eval[[1]]))
```

## Calibration of evaporation soil model output for a verification dataset

The parameters of the previous section now are applied to the data.verification period. This period will be defined from 2017-01-01 to 2017-12-31. Then, the statistics parameters of the calibration in this dataset are shown:
```{r}
data.verification <- data[which(data$Dates == "2017-01-01"):which(data$Dates == "2017-12-31"),]

verif.eval <- mg.evaluation(input.data = data.verification, predictand = 'evapo_obs',
                            predictors = predictors.variables,
                            var.model = 'OUT_EVAP',
                            lmodel = ml.model)

verif.eval[[2]]
```

Finally, the monthly plot of this dataset is displayed below:
```{r}
ploting(daily2monthly(data = verif.eval[[1]]))
```