---
title: Relative humidity calibration for the WRF model
layout: post
subtitle: Gonzalo M. Díaz
leafletmap: true
always_allow_html: yes
last_modified_at: 2024-12-04
output: 
  md_document:
    variant: gfm
    preserve_yaml: true
---

Post for calibration of relative humidity of WRF model

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data download

First, the aws.wrfsmn and others libraries should be open:
```{r, message = FALSE}
library("aws.wrfsmn")
library("dplyr")
library("terra")
library("tibble")
library("ggplot2")
```

Then, some WRF filenames are defined from WRF SMN AWS service
```{r}
filenames <- c()
for (i in c(1, 4))
  {aux.filenames <- get.wrf.files(year = 2023, month = 1, day = i, cycle = 00, time = "01H")
   filenames <- c(filenames, aux.filenames)
  }
print(filenames[1:13])
```

The data of WRF model is downloaded from AWS (<https://registry.opendata.aws/smn-ar-wrf-dataset/>). The documentation of this dataset can be obtained from: <https://odp-aws-smn.github.io/documentation_wrf_det/>.

Finally the dataset is downloaded from the 'filenames' variable:
```{r, eval = FALSE}
wrf.download(wrf.name = filenames)
```

After this line, the dataset defined in 'filenames' will be downloaded.

## Definition of predictors variables

The variables for the adjustment can be any of those found in the WRF dataset. Here would be:
```{r}
variables <- rast("./_includes/WRFDETAR_01H_20230101_00_000.nc") %>% names()
print(variables)
```

The definition of each variable can be obtained from: <https://odp-aws-smn.github.io/documentation_wrf_det/Formato_de_datos/>. In this case the T2, HR2, SMOIS and magViento10 will be the variables to use for the calibration of the Relative Humidity (but other combinations of variables could be a better fit).

The nc files are open with the terra package and separated in each variable:
```{r}
files <- list.files(path = "./_includes/", pattern = "nc", full.names = TRUE)
nc.files <- rast(files)

T2 <- nc.files[[which(names(nc.files) == "T2")]]
HR2 <- nc.files[[which(names(nc.files) == "HR2")]]
SMOIS <- nc.files[[which(names(nc.files) == "SMOIS")]]
magViento10 <- nc.files[[which(names(nc.files) == "magViento10")]]
```

Before extraction of data by location point, the transformation of the reference coordinate system should be done:
```{r, results = "hide"}
T2 <- project(T2, "+proj=longlat +datum=WGS84", method = "bilinear")
HR2 <- project(HR2, "+proj=longlat +datum=WGS84", method = "bilinear")
SMOIS <- project(SMOIS, "+proj=longlat +datum=WGS84", method = "bilinear")
magViento10 <- project(magViento10, "+proj=longlat +datum=WGS84", method = "bilinear")
```

The location in which will try to calibrate the WRF model will be Sunchales City, in the center of Argentina, the location of this city is:
```{r}
LON <- -61.53258
LAT <- -30.95686
```

Now, the temporal series of each variable are taken in that location:
```{r}
T2.ts <- extract(T2, vect(cbind(LON, LAT)), ID = FALSE)
HR2.ts <- extract(HR2, vect(cbind(LON, LAT)), ID = FALSE)
SMOIS.ts <- extract(SMOIS, vect(cbind(LON, LAT)), ID = FALSE)
magViento.ts <- extract(magViento10, vect(cbind(LON, LAT)), ID = FALSE)
```

Now the data is arranged in a table with the Date information as first column:
```{r}
data.table <- tibble(Date = time(T2),
                     T2.wrf = t(T2.ts),
                     HR2.wrf = t(HR2.ts),
                     SMOIS.wrf = t(SMOIS.ts),
                     magViento.wrf = t(magViento.ts))
data.table
```

## Predicting variable: observational data

The variable to predict, will be the relative humidity observed in Sunchales City. Here is a sample of the data:
```{r}
data.obs <- read.table("./data/sunchales.txt", header = TRUE) %>% as_tibble()
data.obs
```

The relative humidity observed can be found in the last column. Now the data.obs is append next to the data.table:
```{r}
data.table <- cbind(data.table, c(data.obs$humrel[1:96], data.obs$humrel[1:50]))
colnames(data.table) <- c("Dates", "T2.wrf", "HR2.wrf", "SMOIS.wrf", "magViento.wrf", "HR2.obs")
```

## Training and verification of the relative humidity calibration

The data now will be trained with the 2023-01-01 00:00:00 to 2023-01-03 23:00:00 period using 'multiple.guidance' function with the predictors variables defined with T2.wrf, HR2.wrf, SMOIS.wrf and magViento.wrf:

```{r}
data.training <- data.table[1:which(data.table$Date == "2023-01-03 23:00:00"),]

ml.model <- multiple.guidance(input.data = data.training,
                              predictand = 'HR2.obs',
                              predictors = c('T2.wrf', 'HR2.wrf', 'SMOIS.wrf', 'magViento.wrf'))
ml.model$coefficients
```

Now, the parameters can be used to evaluate the model in any dataset. Here it is applied to the same training period:
```{r}
train.eval <- mg.evaluation(input.data = data.training,
                            predictand = 'HR2.obs',
                            predictors = c('T2.wrf', 'HR2.wrf', 'SMOIS.wrf', 'magViento.wrf'),
                            var.model = 'HR2.wrf',
                            lmodel = ml.model)
```

The plot can not be visualize with 'ploting' function, because this function only works with daily data, in this case the data is hourly. So, the ggplot function will be use to visualize the results:
```{r}
figure <- ggplot(data = train.eval[[1]]) +
  geom_line(aes(Dates, observation, colour = "red"), group = 1) +
  geom_line(aes(Dates, model, colour = "blue"), group = 1) +
  geom_line(aes(Dates, guidance), group = 1)

print(figure)
```

The second element of the list has the statistics parameters of the calibration: -->
```{r}
train.eval[[2]]
```


<!-- ## Calibration of evaporation soil model output for a verification dataset -->

<!-- The parameters of the previous section now are applied to the data.verification period. This period will be defined from 2017-01-01 to 2017-12-31. Then, the statistics parameters of the calibration in this dataset are shown: -->
<!-- ```{r} -->
<!-- data.verification <- data[which(data$Dates == "2017-01-01"):which(data$Dates == "2017-12-31"),] -->

<!-- verif.eval <- mg.evaluation(input.data = data.verification, predictand = 'evapo_obs', -->
<!--                             predictors = predictors.variables, -->
<!--                             var.model = 'OUT_EVAP', -->
<!--                             lmodel = ml.model) -->

<!-- verif.eval[[2]] -->
<!-- ``` -->

<!-- Finally, the monthly plot of this dataset is displayed below: -->
<!-- ```{r} -->
<!-- ploting(daily2monthly(data = verif.eval[[1]])) -->
<!-- ``` -->