---
title: Calibration: tidymodels vs aws.wrfsmn
layout: post
subtitle: Gonzalo M. Díaz
leafletmap: true
always_allow_html: yes
last_modified_at: 2026-03-28
output: 
  md_document:
    variant: gfm
    preserve_yaml: true
---

Post for calibration of soil model evaporation output from machine learning models of tidymodels package and aws.wrfsmn package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

Library that are going to be used:

```{r, echo=FALSE}
library('tidymodels')
library('glmnet')
library('aws.wrfsmn')
```

The example data to use will be ‘eva’ again and can be visualize:

```{r}
options(width = 200)
head(eva)
```

As a reminder, the dataset contains multiple columns representing different hydrometeorological variables — such as precipitation, evaporation, runoff, and soil moisture, among others. All variables are derived from the VIC (Variable Infiltration Capacity) model (see documentation, <https://vic.readthedocs.io/en/master/Overview/ModelOverview/>), with the exception of the evapo_obs column, which corresponds to observed evaporation data for the same location. The dataset spans from 2015-01-01 to 2017-12-31.

## Definition of predictors variables

The variables chosen for adjustment can be any of those modeled. Here they will be:

```{r}
predictors.variables <- c('OUT_PREC', 'OUT_EVAP', 'OUT_RUNOFF', 'OUT_BASEFLOW',
                          'OUT_SOIL_MOIST_lyr_1', 'OUT_SURF_TEMP')
```

There is no limit on which variables to take, they could be more or less depending on what the user wants.

## Definition of parameters of the Multiple Linear Regression

The data now will be trained with the 2015-01-01 to 2017-12-31 period using 'multiple.guidance' function from the `aws.wrfsmn` package:

```{r}
data <- eva

ml.model <- multiple.guidance(input.data = data,
                              predictand = 'evapo_obs',
                              predictors = predictors.variables)
ml.model$coefficients
```

Now the data will be trained using the **GLMNET model** from the `tidymodels` framework in R.

```{r}
# Define the GLMNET model
glmn_model <- linear_reg(
  penalty = tune(),
  mixture = tune()
) %>%
  set_engine("glmnet") %>%
  set_mode("regression")

# Create a recipe
glmn_recipe <- recipe(evapo_obs ~ ., data = data) %>%
  update_role(Dates, new_role = "ID") %>%
  step_naomit(all_predictors(), all_outcomes()) %>%
  step_normalize(all_predictors())

# Create a workflow
glmn_workflow <- workflow() %>%
  add_recipe(glmn_recipe) %>%
  add_model(glmn_model)

# Define resampling strategy (e.g., cross-validation)
cv_folds <- vfold_cv(data, v = 5)

# Tune hyperparameters
glmn_tune <- tune_grid(
  glmn_workflow,
  resamples = cv_folds,
  grid = 20
)

# Select the best model
best_glmn <- select_best(glmn_tune, metric = "rmse")

# Finalize the workflow
final_glmn <- finalize_workflow(glmn_workflow, best_glmn)

# Fit the final model on the data
glmn_fit <- final_glmn %>%
  fit(data = data)

# Evaluate on test data
glmn_results <- glmn_fit %>%
  predict(data) %>%
  bind_cols(data) %>%
  metrics(truth = evapo_obs, estimate = .pred)

print(glmn_results)
```

## Calibration of evaporation soil model output for a verification dataset

The parameters of the previous section now are applied to the data.verification period. This period will be defined from 2017-01-01 to 2017-12-31. Then, the statistics parameters of the calibration in this dataset are shown:

```{r}
data.verification <- data[which(data$Dates == "2017-01-01"):which(data$Dates == "2017-12-31"),]

verif.eval <- mg.evaluation(input.data = data.verification, predictand = 'evapo_obs',
                            predictors = predictors.variables,
                            var.model = 'OUT_EVAP',
                            lmodel = ml.model)

verif.eval[[2]]
```

Finally, the monthly plot of this dataset is displayed below:

```{r}
ploting(daily2monthly(data = verif.eval[[1]]))
```
