---
title: Calibration: tidymodels vs aws.wrfsmn
layout: post
subtitle: Gonzalo M. Díaz
leafletmap: true
always_allow_html: yes
last_modified_at: 2026-03-28
output: 
  md_document:
    variant: gfm
    preserve_yaml: true
---

Post for calibration of soil model evaporation output from machine learning models of tidymodels package and aws.wrfsmn package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

Library that are going to be used:

```{r, echo=FALSE}
library('tidymodels')
library('glmnet')
library('aws.wrfsmn')
library('zoo')
library('hydroTSM')
```

The example data to use will be ‘eva’ again and can be visualize:

```{r}
options(width = 200)
head(eva)
```

As a reminder, the dataset contains multiple columns representing different hydrometeorological variables — such as precipitation, evaporation, runoff, and soil moisture, among others. All variables are derived from the VIC (Variable Infiltration Capacity) model (see documentation, <https://vic.readthedocs.io/en/master/Overview/ModelOverview/>), with the exception of the evapo_obs column, which corresponds to observed evaporation data for the same location. The dataset spans from 2015-01-01 to 2017-12-31.

## Definition of predictors variables

The variables chosen for adjustment can be any of those modeled. Here they will be:

```{r}
predictors.variables <- c('OUT_PREC', 'OUT_EVAP', 'OUT_RUNOFF', 'OUT_BASEFLOW',
                          'OUT_SOIL_MOIST_lyr_1', 'OUT_SURF_TEMP')
```

There is no limit on which variables to take, they could be more or less depending on what the user wants.

## Definition of parameters and models

The data now will be trained with the 2015-01-01 to 2017-12-31 period using 'multiple.guidance' function from the `aws.wrfsmn` package:

```{r}
data <- eva

ml.model <- multiple.guidance(input.data = data,
                              predictand = 'evapo_obs',
                              predictors = predictors.variables)
ml.model$coefficients

verif.eval <- mg.evaluation(input.data = data, predictand = 'evapo_obs',
                            predictors = predictors.variables,
                            var.model = 'OUT_EVAP',
                            lmodel = ml.model)

verif.eval[[2]]
```

Now the data will be trained using the **GLMNET model** from the `tidymodels` framework in R.

```{r}
# Define the GLMNET model
glmn_model <- linear_reg(
  penalty = tune(),
  mixture = tune()
) %>%
  set_engine("glmnet") %>%
  set_mode("regression")

# Create a recipe
glmn_recipe <- recipe(evapo_obs ~ ., data = data) %>%
  update_role(Dates, new_role = "ID") %>%
  step_naomit(all_predictors(), all_outcomes()) %>%
  step_normalize(all_predictors())

# Create a workflow
glmn_workflow <- workflow() %>%
  add_recipe(glmn_recipe) %>%
  add_model(glmn_model)

# Define resampling strategy (e.g., cross-validation)
cv_folds <- vfold_cv(data, v = 5)

# Tune hyperparameters
glmn_tune <- tune_grid(
  glmn_workflow,
  resamples = cv_folds,
  grid = 20
)

# Select the best model
best_glmn <- select_best(glmn_tune, metric = "rmse")

# Finalize the workflow
final_glmn <- finalize_workflow(glmn_workflow, best_glmn)

# Fit the final model on the data
glmn_fit <- final_glmn %>%
  fit(data = data)

# Evaluate on test data
glmn_results <- glmn_fit %>%
  predict(data) %>%
  bind_cols(data) %>%
  metrics(truth = evapo_obs, estimate = .pred)

print(glmn_results)
```

## Plots of models

Now the results of the model ran with `aws.wrfsmn` are taken and transformed to monthly data. Note that the daily2monthly used here is from the `aws.wrfsmn` package:

```{r}
model.awswrfsmn <- mg.evaluation(input.data = data,
                                 predictand = 'evapo_obs',
                                 predictors = predictors.variables,
                                 var.model = 'OUT_EVAP',
                                 lmodel = ml.model)

multiple.monthly <- aws.wrfsmn::daily2monthly(model.awswrfsmn[[1]])
```

The prediction of the GLMNET model also is transformed to monthly data with hydroTSM and zoo libraries. Note that the daily2monthly used here is from the `hydroTSM` package:

```{r}
model.glmnet <- glmn_fit %>% predict(data)

# Create data frame
model.glmnet <- data.frame(Dates = data$Dates, glmnet = model.glmnet)

# Transform to zoo type
glmnet.zoo <- zoo(model.glmnet$.pred, order.by = as.Date(model.glmnet$Dates))

# Aggregate to monthly
glmnet.monthly <- hydroTSM::daily2monthly(glmnet.zoo, FUN = mean)
```

Unify the data in one and only data frame

```{r}
data.to.print <- cbind(multiple.monthly, glmnet.monthly)
head(data.to.print)
```

Finally, the monthly plot of the models are displayed below:

```{r}

ggplot(data.to.print) +
  geom_line(aes(Dates, observation)) +
  geom_line(aes(Dates, model)) +
  geom_line(aes(Dates, guidance))
  #geom_line(aes(Dates, glmnet.monthly))

```
