---
title: Evaporation soil model calibration
subtitle: Gonzalo M. Díaz
leafletmap: true
always_allow_html: yes
last_modified_at: 2024-11-10
output: 
  md_document:
    variant: gfm
    preserve_yaml: true
---

Post for calibration of soil model evaporation output


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

First, the aws.wrfsmn library should be open:

```{r}
library("aws.wrfsmn")
```

The example data to use will be ‘eva’ and can be visualize:

```{r}
options(width = 200)
head(eva)
```

The data has several columns with different hydrometeorological variables, for example: precipitation, evaporation, runoff, soil moisture, etc. All these variables are obtained from the VIC (Variable Infiltration Capacity, <https://vic.readthedocs.io/en/master/Overview/ModelOverview/>) soil model, except for the 'evapo_obs' column, which represents the observed data of evaporation for the same location. The data goes from 2015-01-01 to 2017-12-31.

## Definition of predictors variables

The variables chosen for adjustment can be any of those modeled. Here they will be:

```{r}
predictors.variables <- c('OUT_PREC', 'OUT_EVAP', 'OUT_RUNOFF', 'OUT_BASEFLOW',
                          'OUT_SOIL_MOIST_lyr_1', 'OUT_SURF_TEMP')
```

There is no limit on which variables to take, they could be more or less depending on what the user wants.
