---
title: Calibration - tidymodels vs aws.wrfsmn
layout: post
subtitle: Gonzalo M. Díaz
leafletmap: true
always_allow_html: yes
last_modified_at: 2026-03-31
output: 
  md_document:
    variant: gfm
    preserve_yaml: true
---

Post for calibration of soil model evaporation output from machine learning models of tidymodels package and aws.wrfsmn package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

Library that are going to be used:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library('tidymodels')
library('knitr')
library('glmnet')
library('aws.wrfsmn')
library('zoo')
library('hydroTSM')
```

The example data to use will be ‘eva’ again and can be visualize:

```{r}
options(width = 200)
head(eva)
```

As a reminder, the dataset contains multiple columns representing different hydrometeorological variables — such as precipitation, evaporation, runoff, and soil moisture, among others. All variables are derived from the VIC (Variable Infiltration Capacity) model (see documentation, <https://vic.readthedocs.io/en/master/Overview/ModelOverview/>), with the exception of the evapo_obs column, which corresponds to observed evaporation data for the same location. The dataset spans from 2015-01-01 to 2017-12-31.

## Definition of predictors variables

The variables chosen for adjustment can be any of those modeled. Here they will be:

```{r}
predictors.variables <- c('OUT_PREC', 'OUT_EVAP', 'OUT_RUNOFF', 'OUT_BASEFLOW',
                          'OUT_SOIL_MOIST_lyr_1', 'OUT_SURF_TEMP')
```

There is no limit on which variables to take, they could be more or less depending on what the user wants.

## Definition of models

The data now will be trained with the 2015-01-01 to 2017-12-31 period using 'multiple.guidance' function from the `aws.wrfsmn` package. This implementation relies on a multivariate linear regression model to predict the output variable.

```{r}
data <- eva

mlr.model <- multiple.guidance(input.data = data,
                               predictand = 'evapo_obs',
                               predictors = predictors.variables)

verif.eval <- mg.evaluation(input.data = data, predictand = 'evapo_obs',
                            predictors = predictors.variables,
                            var.model = 'OUT_EVAP',
                            lmodel = mlr.model)
```

On the other hand, the `parsnip` package provides a complete list of all models available within the `tidymodels` ecosystem.

```{r}
models <- get_from_env("models")

kable(data.frame(Models = models), caption = "Available models in parsnip")
```

From this list, we can select the model that best suits our specific problem or application. In this case, we will use a **Support Vector Machine (SVM)** polinomial model (svm_poly). Now the data will be trained using the **SVM model** from the `tidymodels` framework in R.

```{r}
# Define the model from table
svm_poly_kernlab_spec <- svm_poly(cost = tune(), degree = tune(),
                                  scale_factor = tune(), margin = tune()) %>%
  set_engine('kernlab') %>%
  set_mode('regression')

# Create a recipe
svm_recipe <- recipe(evapo_obs ~ ., data = data) %>%
  update_role(Dates, new_role = "ID") %>%
  step_naomit(all_predictors(), all_outcomes()) %>%
  step_normalize(all_predictors())

# Create a workflow
svm_workflow <- workflow() %>%
  add_recipe(svm_recipe) %>%
  add_model(svm_poly_kernlab_spec)

# Define resampling strategy (e.g., cross-validation)
cv_folds <- vfold_cv(data, v = 5)

# Tune hyperparameters
svm_tune <- tune_grid(
  svm_workflow,
  resamples = cv_folds,
  grid = 20
)

# Select the best model
best_svm <- select_best(svm_tune, metric = "rmse")

# Finalize the workflow
final_svm <- finalize_workflow(svm_workflow, best_svm)

# Fit the final model on the data
svm_fit <- final_svm %>%
  fit(data = data)

# Evaluate on test data
svm_results <- svm_fit %>%
  predict(data) %>%
  bind_cols(data) %>%
  metrics(truth = evapo_obs, estimate = .pred)
```

In the next section some of the svm_results will be shown.

## Plots of models

Now the results of the model ran with `aws.wrfsmn` are taken and transformed to monthly data. Note that the daily2monthly used here is from the `aws.wrfsmn` package:

```{r}
model.awswrfsmn <- mg.evaluation(input.data = data,
                                 predictand = 'evapo_obs',
                                 predictors = predictors.variables,
                                 var.model = 'OUT_EVAP',
                                 lmodel = mlr.model)

multiple.monthly <- aws.wrfsmn::daily2monthly(model.awswrfsmn[[1]])
```

The prediction of the SVM model also is transformed to monthly data with `hydroTSM` and `zoo` libraries. Note that the daily2monthly used here is from the `hydroTSM` package:

```{r}
model.svm <- svm_fit %>% predict(data)

# Create data frame
model.svm <- data.frame(Dates = data$Dates, glmnet = model.svm)

# Transform to zoo type
svm.zoo <- zoo(model.svm$.pred, order.by = as.Date(model.svm$Dates))

# Aggregate to monthly
svm.monthly <- hydroTSM::daily2monthly(svm.zoo, FUN = mean)
```

Unify the data in one and only data frame

```{r}
data.to.print <- cbind(multiple.monthly, svm.monthly)
head(data.to.print)
```

Finally, the monthly plot of the models are displayed below:

```{r}
ggplot(data.to.print) +
  geom_line(aes(Dates, observation, color = "Observed")) +
  geom_line(aes(Dates, model, color = "VIC Model")) +
  geom_line(aes(Dates, guidance, color = "MLR Model")) +
  geom_line(aes(Dates, svm.monthly, color = "SVM Poly Model")) +
  scale_color_manual(values = c(
    "Observed"  = "black",
    "VIC Model"     = "steelblue",
    "MLR Model"  = "tomato",
    "SVM Poly Model"    = "seagreen"
  )) +
  labs(color = "Legend") +
  theme_minimal()

```

From the plot, we can observe that the model that best represents the temporal evolution of observed evaporation is the multivariate linear model from the `aws.wrfsmn` package, primarily due to its improved representation of peak values. Nevertheless, the SVM model also produces an improvement in the representation of evaporation compared to the raw output of the VIC model (blue line). It is essential to apply bias corrections to the raw outputs of mathematical models in order to achieve a better representation of reality.

Finally, as a closing analysis, the Root Mean Square Error (RMSE) for each of the three models is presented below:

```{r}
kable(data.frame(RMSE.VIC = verif.eval[[2]]$rmse[1],
                 RMSE.MLR = verif.eval[[2]]$rmse[2],
                 RMSE.SVM = svm_results$.estimate[1]))
```
